---
title: "ESM 211 Applied Population Ecology Final Paper"
subtitle: "Fancy title about Skipjack, including scientific name and shit"
author: "Villaseñor-Derbez, J.C., Faro, C"
output:
  pdf_document: default
---

\clearpage

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
suppressPackageStartupMessages({
  library(tidyverse)
  library(stargazer)
  library(popbio)
  library(knitr)
  library(ggExtra)
})
```

## Define the functions we will need

### Age at length

$$A = \frac {1}{-K} \times log(1-(\frac {L}{L_{inf}}))+t_0$$

Where:

  - $A$ = age (years)
  - $K$ = growth rate
  - $L$ = length (cm)
  - $L_{inf}$ = asymptotic length (cm)
  
The model can then be made into a function as follows:

```{r}
# Convert length to age using von bertalanffy model, solving for A
length2age <- function(length, l_inf, K, t_o){
  age <- (1/-K)*(log(1-(length/L_inf))) + t_o
  return(age)
}
```

### Length at age

Solving for $L$ in the equation above, we obtain:

$$L = l_{inf}(1-exp^{(-K(A-t_o))})$$

Which can be made into a function as:

```{r}
# Convert age to length using von bertalanffy model
age2length <- function(age, l_inf, K, t_o){
  length <- l_inf*(1-exp(-K*(age-t_o)))
  return(length)
}
```

## Fecundity at length

$$f = 10^{(a+(b*log_{10}(L*10)))}$$

Where:

  - $f$ = fecundity (number of eggs)
  - $a$ = **NAME HERE**
  - $b$ = **NAME HERE**
  - $L$ = length (cm)
  
As a function, this takes the form:

```{r}
#Convert length to fecundity (number of eggs)
fecundity <- function(length, a, b){
  f <- 10^(a+(b*log10(length*10)))
  return(f)
}
```

## Define demographic parameters

```{r}
# Or from Us and Stiurskjgfas, 1981
L_inf <- 102
K <- 0.55
t_0 <- -0.02

# Fecundity parameters
fec_a <- -1.33354
fec_b <- 3.238

# Mortality
m <- 0.63
z <- 1.39

```


```{r, results='asis'}
A <- matrix(0, 13, 13) #initial empty matrix with all 0

# Populate matrix with mortality
for (i in 2:13){
  A[i,i-1] <- exp(-z)
}

# Populate matrix with fecundity
ages <- seq(1,13)
lengths <- age2length(ages, L_inf, K, t_0)
A[1,] <- fecundity(lengths, fec_a, fec_b) * 0.5 * 1e-7 #Multiply fecundity times 0.5 (F:M = 1:1) and times 1e-7 (survival of 1 year olds)

colnames(A) <- ages
rownames(A) <- ages

kable(A,
      digits = 2,
      col.names = paste0("$a_{",ages,"}$"),
      row.names = F,
      caption = "Population matrix A. The inferior diagonal represents survivals, while the first row represents facundities.")

```


## Build a population vector

```{r, warning = F, fig.cap="Initial population population structure, represented by ages (x-axis) and relative abundances (y-axis). Numbers above bars indicate the corresponding size (TL cm) for each age."}

load("size_dist_1980.RData") #This is the n-vector for 1980, loads it to a vector called n

n <- n %>% {
  .$N}%>% 
  c(rep(0, times = 8))

data.frame(N = n,
           Age = ages) %>% 
  mutate(N = N/sum(N)) %>% 
  ggplot(aes(x = Age, y = N)) +
  geom_bar(stat = "identity", fill = "gray", color = "black", size = 1) +
  geom_text(aes(x = Age, y = N + 0.05, label = formatC(lengths, digits = 0, format = "f"))) +
  theme_bw() +
  scale_x_continuous(expand = c(0.01, 0.01), breaks = seq(1, 13)) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1), labels = scales::percent) +
  removeGridX() +
  theme(text = element_text(size = 14)) +
  labs(x = "Age (years)", y = "log-transformed abundance")

```

```{r}
project <- popbio::pop.projection(A, n, 20)
```

```{r, fig.cap = "Figure 1 - Population size through time, represented by ages."}
project$stage.vectors %>% 
  as.data.frame() %>% 
  mutate(Age = as.factor(seq(1,13))) %>% 
  gather(Year, N, -Age) %>% 
  mutate(Year = as.numeric(as.character(Year))) %>% 
  select(Year, Age, N) %>% 
  ggplot(aes(x = Year, y = log(N), color = Age)) +
  geom_line() +
  theme_bw()

```

```{r}

A_min <- A

A_min[2, 1] <- exp(-m)

knitr::kable(A_min,
      digits = 2,
      col.names = paste0("$a_{",ages,"}$"),
      row.names = F,
      caption = "Population matrix $\\hat{A}$, modifying mortality of age 1 organisms through a minimum catch size set at 67 cm. The inferior diagonal represents survivals, while the first row represents facundities.")
  
```

```{r}
project_min <- popbio::pop.projection(A_min, n, 20)
```

```{r, fig.cap = "Population size through time, represented by ages."}
project_min$stage.vectors %>% 
  as.data.frame() %>% 
  mutate(Age = as.factor(seq(1, 13))) %>% 
  gather(Year, N, -Age) %>% 
  mutate(Year = as.numeric(as.character(Year))) %>% 
  select(Year, Age, N) %>% 
  ggplot(aes(x = Year, y = log(N), color = Age)) +
  geom_line() +
  theme_bw()
```

```{r, fig.cap= "Total population size through time."}

tot1 <- project$pop.sizes
tot2 <- project_min$pop.sizes
time <- seq(1:20)

data.frame(time, BAU = tot1, Int = tot2) %>% 
  gather(scenario, popsize, -time) %>% 
  ggplot(aes(x = time, y = log(popsize), color = scenario)) +
  geom_point(size = 2) +
  geom_line(size = 1) +
  theme_bw() +
  scale_color_brewer(palette = "Set1")

```

```{r, fig.cap = "Stable stage population structure, represented by ages (x-axis) and relative abundances (y-axis). Numbers above bars indicate the corresponding size (TL cm) for each age."}
data.frame(Age = ages, BAU = project$stable.stage, Int = project_min$stable.stage) %>% 
  gather(scenario, N, -Age) %>% 
  ggplot(aes(x = Age, y = N)) +
  geom_bar(stat = "identity", fill = "gray", color = "black", size = 1) +
  facet_grid(~scenario, scales = "free_y") +
  geom_text(aes(x = Age, y = N + 0.05, label = formatC(age2length(Age,  L_inf, K, t_0), digits = 0, format = "f")), size = 3) +
  theme_bw() +
  scale_x_continuous(expand = c(0.01, 0.01), breaks = seq(1, 13)) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1), labels = scales::percent) +
  removeGridX() +
  theme(text = element_text(size = 14)) +
  labs(x = "Age (years)", y = "Relative abundance")
  
```


```{r}
data.frame(Age = ages, BAU = project$stable.stage, Int = project_min$stable.stage) %>% 
  mutate(Diff = Int - BAU) %>% 
  ggplot(aes(x = Age, y = Diff)) +
  geom_bar(stat = "identity", fill = "gray", color = "black", size = 1) +
  theme_bw() +
  scale_x_continuous(expand = c(0.01, 0.01), breaks = seq(1, 13)) +
  scale_y_continuous(expand = c(0.01, 0.01), labels = scales::percent) +
  removeGridX() +
  theme(text = element_text(size = 14)) +
  labs(x = "Age (years)", y = "Change in relative abundance")
```

